```{r set}
#| include: false
library(tidyverse)
library(sf)
library(terra)
```

# Grid {.unnumbered}

We aim at using a 10 km by 10 km grid, but will first do a test with a 20 km by 20 km grid. We will first do a test with 400 simulations, covering an area of 20 x 20 = 400 km or approximatively 0.2°. We will use cell centre to define environment and simulate a representative 4-ha simulation. We can first represent the grid and evaluate expected biomass variation from @fayad2016 and soil and climate variation from input data (see in dedicated pages). **The issue is that there is low biomass variation in the area in which to conduct the test. Perhaps we should use two areas.**

```{r grid}
#| message: false
#| warning: false
#| fig-cap: "Caption."
guyafor <- read_tsv("data/raw_data/guyafor.tsv") %>% 
  filter(Forest %in% c("Nouragues", "Montagne Tortue")) %>% 
  select(Forest, Plot, Xutm, Yutm) %>% 
  group_by(Forest, Plot) %>% 
  summarise_all(mean, na.rm = TRUE) %>% 
  na.omit() %>% 
  st_as_sf(coords = c("Xutm", "Yutm"), crs = 32622)
area <- tibble(lon = c(-52.7, -52.3), lat = c(4.0, 4.4)) %>% 
  st_as_sf(coords = c("lon", "lat"), crs = 4326) %>% 
  st_bbox() %>% 
  st_as_sfc()
grid <- st_make_grid(area, n = c(20, 20))
agb <- rast("data/raw_data/fayad_2016/AGB_16122015.tif") %>% 
  crop(area %>% st_transform(crs = 32622) %>% st_buffer(2*10^3))
ggplot(grid) +
  tidyterra::geom_spatraster(data = agb, alpha = .5) +
  scale_fill_viridis_c(na.value = "transparent", guide = "none") +
  geom_sf(fill = NA) +
  geom_sf(data = guyafor) +
  theme_bw()
```

The within cell variation in biomass was low with a median around 10 Mg/ha.

```{r delta_agb}
#| message: false
#| warning: false
#| fig-cap: "Caption."
extract(agb, vect(grid%>% st_transform(crs = 32622)), exact = TRUE) %>% 
  group_by(ID) %>% 
  mutate(delta = abs(AGB_16122015 - mean(AGB_16122015))) %>% 
  ggplot(aes(delta)) +
  geom_histogram() +
  theme_bw() +
  xlab(expression('Cell'~Delta~AGB~"["~Mg~ha~"]"))
```
