```{r set}
#| include: false
library(tidyverse)
library(sf)
library(terra)
```

# Grid {.unnumbered}

We decided to use a 10 km grid, first testing on a total area of 0.1°, corresponding to 90 simulations after removing empty cell (more than 50% missing value). We will use cell centre to define environment and simulate a representative 4-ha simulation. We can first represent the grid and evaluate expected biomass variation from @fayad2016 and soil and climate variation from input data (see in dedicated pages).

```{r grid}
#| message: false
#| warning: false
#| fig-cap: "Caption."
area <- tibble(lon = c(-53, -52), lat = c(4.0, 5.0)) %>% 
  st_as_sf(coords = c("lon", "lat"), crs = 4326) %>% 
  st_bbox() %>% 
  st_as_sfc()
guyafor <- read_tsv("data/raw_data/guyafor.tsv") %>% 
  select(Forest, Plot, Xutm, Yutm) %>% 
  group_by(Forest, Plot) %>% 
  summarise_all(mean, na.rm = TRUE) %>% 
  na.omit() %>% 
  st_as_sf(coords = c("Xutm", "Yutm"), crs = 32622) %>% 
  st_intersection(area %>% st_transform(crs = 32622))
grid <- st_make_grid(area, n = c(10, 10))
agb <- rast("data/raw_data/fayad_2016/AGB_16122015.tif") %>% 
  crop(area %>% st_transform(crs = 32622) %>% st_buffer(2*10^3))
non_empty <- extract(agb, vect(grid%>% st_transform(crs = 32622)), exact = TRUE) %>% 
  mutate(na = is.na(AGB_16122015)) %>% 
  group_by(ID) %>% 
  summarise(na = sum(na)/n()) %>% 
  filter(na <= 0.5)
grid <- grid[non_empty$ID]
ggplot(grid) +
  tidyterra::geom_spatraster(data = agb, alpha = .5) +
  scale_fill_viridis_c(na.value = "transparent", guide = "none") +
  geom_sf(fill = NA) +
  geom_sf(data = guyafor) +
  theme_bw()
```

The within cell variation in biomass was low with a median around 13.5 Mg/ha.

```{r delta_agb}
#| message: false
#| warning: false
#| fig-cap: "Caption."
extract(agb, vect(grid%>% st_transform(crs = 32622)), exact = TRUE) %>% 
  group_by(ID) %>% 
  mutate(delta = abs(AGB_16122015 - mean(AGB_16122015))) %>% 
  # ungroup() %>% summarise_all(median, na.rm = TRUE)
  ggplot(aes(delta)) +
  geom_histogram() +
  theme_bw() +
  xlab(expression('Cell'~Delta~AGB~"["~Mg~ha~"]"))
```

```{r grid_save}
#| eval: false
grid %>% 
  st_centroid() %>% 
  st_coordinates() %>% 
  as_tibble() %>% 
  mutate_all(round, 2) %>% 
  write_tsv("data/derived_data/grid.tsv")
```
